---
title: "Cluster Number Analysis"
output: html_document
date: '2022-07-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())
gc()
library(tidyverse)  # data manipulation
library(ggplot2)
library(psych)
library(analogue)
library(data.table)
library(trqwe)
library(DiffCorr)
library(clue)
library(flowAssist)
library(FlowSOM)
library(FuseSOM)
library(diptest)
set.seed(1994)
source.all('/home/ewillie/Documents/PhD/Papers/FuseSOM_Paper/helper_scripts')
```

# Maldegem2021
## read in the data and get markers of interest
```{r}
maldegem_dat <- as.data.frame(fread(
  '~/Documents/PhD/Spatial_Clustering/Maldegem2021_LungCancer_IMC/Maldegem2021_processed.csv'))

# remove the unclassified cell types
# maldegem_dat <- maldegem_dat[maldegem_dat$clustername3 != 'Unclassified', ]

# rename the cluster column
names(maldegem_dat)[names(maldegem_dat) == 'clustername3'] <- 'CellType'

maldegem_markers <- colnames(maldegem_dat)[1:17]

# generate subsets of the data
# maldegem_subsets <- subsetMatrix(maldegem_dat)

maldegem_res <- runClustNum(maldegem_dat, maldegem_markers)

rm(maldegem_dat)
gc()
```


# Rendeiro2020
## Read in the data and set markers
```{r}
rendeiro_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_intensities.csv',
             nThread = 7))

# get the metadata
rendeiro_meta <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_metadata.csv',
             nThread = 7))

# extract the markers of interest and add the cluster labels and image metatdata
rendeiro_markers <- colnames(rendeiro_dat)[1:38]
rendeiro_dat <- rendeiro_dat[, rendeiro_markers]

# combine the metadata with the intensities and umap coordinates
rendeiro_dat <- cbind(rendeiro_dat, cbind(rendeiro_meta))

names(rendeiro_dat)[names(rendeiro_dat) == 'metacluster_label'] <- 'CellType'
rendeiro_dat <- rendeiro_dat[rendeiro_dat$CellType != "nan", ]

rendeiro_res <- runClustNum(rendeiro_dat, rendeiro_markers)
rm(rendeiro_dat)
gc()
```


# Risom2022
## Read in data and set markers
```{r}
risom_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Risom2022_BreastCancer_IMC/Single_Cell_Data.csv')
)

risom_markers <- c('CD45','SMA','CK7','CK5','VIM','CD31','PanKRT','ECAD',
                   'Tryptase','MPO','CD20','CD3','CD8','CD4','CD14','CD68','FAP',
                   'CD36','CD11c','HLADRDPDQ','P63','CD44')

names(risom_dat)[names(risom_dat) == 'manual_gating_cellineage'] <- 'CellType'

risom_res <- runClustNum(risom_dat, risom_markers)

rm(risom_dat)
gc()
```


# Damond2019 
# Read in data and set markers
```{r}
damond_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Damond2019_Diabetes_IMC/Damond2010_Diabetes_IMC_celldata.csv')
)


# get the markers of interest
damond_markers <- colnames(damond_dat)
damond_markers <- damond_markers[1:36]

# # normalize and scale the data
# damond_dat[, damond_markers] <- None(damond_dat[, damond_markers])
# damond_dat[, damond_markers] <- scale(damond_dat[, damond_markers])

damond_res <- runClustNum(damond_dat, damond_markers)

rm(damond_dat)
gc()
```


# Moldoveanu2022
## read in data and set markers
```{r}
moldoveanu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/ICI_meanIntensity_processed.csv')
)

moldoveanu_dat_non_ICI <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/non-ICI_meanIntensity_processed.csv')
)


moldoveanu_metadata <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/ST4_cell_data.txt')
)


# combine both datasets
moldoveanu_dat <- rbind(moldoveanu_dat, moldoveanu_dat_non_ICI)

# order by object ID
moldoveanu_dat <- moldoveanu_dat[match(moldoveanu_metadata$obj.id, moldoveanu_dat$obj.id), ]

# get markers
moldoveanu_markers <- colnames(moldoveanu_dat)[1:12]

# add celltypes
moldoveanu_dat$CellType <- moldoveanu_metadata$Cluster

moldoveanu_res <- runClustNum(moldoveanu_dat, moldoveanu_markers)

rm(moldoveanu_dat)
gc()
```


# Hoch 2022
## read in data and set markers
```{r}
# read in the data
hoch_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Hoch2022_Melanoma_IMC/hoch2022_Melanoma_counts.csv', 
                  nThread = 7)
)
# get markers of interest
hoch_markers <- colnames(hoch_dat)[1:41]

hoch_res <- runClustNum(hoch_dat, hoch_markers)

rm(hoch_dat)
gc()
```


# Keren2018
## Read in data set markers
```{r}
keren_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Keren2018_BreastCancer_IMC/cellData.csv'))

# set markers
keren_markers <- c('CD45', 'FoxP3', 'CD4', 'CD8', 'CD3', 'CD20', 'CD16', 'CD68', 
                   'MPO', 'HLA-DR', 'Pan-Keratin', 'Keratin17', 'Keratin6', 'p53', 
                   'Beta catenin', 'EGFR')
# set the cell types
names(keren_dat)[names(keren_dat) == 'Group'] <- 'CellType'

keren_res <- runClustNum(keren_dat, keren_markers)

rm(keren_dat)
gc()
```


# McCaffery2022
# load data and get markers
```{r}
# read in the data
mccaferry_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/McCaffery2022_Tuberculosis_IMC/allTB-ME_annotated.csv')
)
# get the markers
mccaferry_markers <- colnames(mccaferry_dat)[6:42]

# set the cell types
names(mccaferry_dat)[names(mccaferry_dat) == 'cell_type'] <- 'CellType'

# remove rows with NAs
mccaferry_dat <- na.omit(mccaferry_dat)

mccaferry_res <- runClustNum(mccaferry_dat, mccaferry_markers)

rm(mccaferry_dat)
gc()
```



## Liu2022
# Get data and extract markers
```{r}
liu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Liu2022_Benchmarking_IMC/cell_table_size_normalized_clusters.csv')
)

liu_markers <- c('CD3', 'CD4', 'CD8', 'CD11c', 'CD20', 'CD31', 'CD45', 'CD56', 
                 'CD68', 'PANCK', 'PAX5','Vimentin')

# define cell types

names(liu_dat)[names(liu_dat) == 'phenotype'] <- 'CellType'

liu_res <- runClustNum(liu_dat, liu_markers)

rm(liu_dat)
gc()
```

## Schurch2020
# Get data and extract markers
```{r}
schurch_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/CODEX_Colon_Schurch2020/CRC_clusters_neighborhoods_markers.csv") %>%
    dplyr::select(-1) %>% 
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(schurch_dat$cluster_name)
schurch_dat$CellType <- factor(schurch_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

# get the markers
schurch_markers <- colnames(schurch_dat)[12:60]

schurch_res <- runClustNum(schurch_dat, schurch_markers)

rm(schurch_dat)
gc()
```

## Phillips2021
# Get data and extract markers
```{r}
phillips_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/Phillips2021_Lymphoma_CODEX/Raw_df_CODEX.csv") %>%
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(phillips_dat$cluster_name)
phillips_dat$CellType <- factor(phillips_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

phillips_dat <- phillips_dat[ , -which(names(phillips_dat) %in% 
                                         c('cd11b','cd16','cd164','ccr4','ccr6',
                                      'egfr', 'p53'))]


# get the markers
phillips_markers <- colnames(phillips_dat)[10:61]

phillips_res <- runClustNum(phillips_dat, phillips_markers)

rm(phillips_dat)
gc()
```

# Bortolomeazzi2020
# Get data and extract markers
```{r}
bortolomeazzi_dat <- fread('~/Documents/PhD/Spatial_Clustering/IMC_ColorectalCancer_Bortolomeazzi2020/Single_Cell_Data.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
bortolomeazzi_markers <- colnames(bortolomeazzi_dat)[16:45]

# add the celltype information
names(bortolomeazzi_dat)[names(bortolomeazzi_dat) == 'Population'] <- 'CellType'


bortolomeazzi_res <- runClustNum(bortolomeazzi_dat, bortolomeazzi_markers)

rm(bortolomeazzi_dat)
gc()
```

# Brbic2022 - Intestine
# Get data and extract markers
```{r}
brbic_int_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/B004_training_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_int_markers <- colnames(brbic_int_dat)[2:49]

# add the celltype information
names(brbic_int_dat)[names(brbic_int_dat) == 'cell_type_A'] <- 'CellType'

brbic_int_res <- runClustNum(brbic_int_dat, brbic_int_markers)

rm(brbic_int_dat)
gc()
```

# Brbic2022 - Tonsil
# Get data and extract markers
```{r}
brbic_tons_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/BE_Tonsil_l3_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_tons_markers <- colnames(brbic_tons_dat)[2:45]

# add the celltype information
names(brbic_tons_dat)[names(brbic_tons_dat) == 'cell_type'] <- 'CellType'

brbic_tons_res <- runClustNum(brbic_tons_dat, brbic_tons_markers)

rm(brbic_tons_dat)
gc()
```

```{r}
res <- rbind(maldegem_res, rendeiro_res, risom_res, damond_res, 
             moldoveanu_res, hoch_res, liu_res, keren_res,
             mccaferry_res, schurch_res, phillips_res,
             bortolomeazzi_res, brbic_int_res, brbic_tons_res) %>% as.data.frame()

res$Dataset <- rep(c('Maldegem et al', 'Rendeiro et al', 'Risom et al', 'Damond et al',
                     'Moldoveanu et al', 'Hoch et al', 'Liu et al',
      'Keren et al', 'Mccaferry et al', 'Schurch et al', 'Phillips et al',
      'Bortolomeazzi et al', 'Brbic et al intestine',
      'Brbic et al tonsil'), each=6)

res$Method <- rep(c("Discrimant","Gap","Slope","Jump","WCD","Silhouette"), 
                  length(unique(res$Dataset)))

# write.csv(res, '~/Documents/PhD/Spatial_Clustering/clustering/clust_est_analysis.csv', row.names = F)

```

Generate Figure 7A
```{r}
res <- read.csv('~/Documents/PhD/Spatial_Clustering/clustering/clust_est_analysis.csv')
colnames(res)[[3]] <- "FM-Index"

res.re <- res %>% dplyr::select(Method, RE, Dataset) %>% as.data.frame() %>% 
  melt(id.var = c('Dataset','Method'))
colnames(res.re) <- c('Dataset', 'Method', 'RE' ,'Relative-Error')
# res.melted$`Relative Error` <- as.numeric(res.melted$`Relative Error`)

p.RE <- ggpaired(res.re, x= 'Method', y = '`Relative-Error`',
                 fill = 'Method', add = 'jitter', palette = 'jama',
                 point.size = 3) + 
  labs(title="",
        x ="Cluster Estimation Method", y = "Relative Error") + theme_bw() +
  theme(
      plot.title = element_text(color="Black", size=25, hjust = 0.5),
      plot.subtitle = element_text(color = "red", size = 25, face = "bold", hjust = 0.5),
      axis.title.x = element_text(color="Black", size=25),
      axis.title.y = element_text(color="Black", size=25),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 25),
      axis.text.y = element_text(size = 25),
      legend.position = "none"
      
    ) +
  geom_hline(yintercept=0, linetype='dotted', col = 'red', linewidth = 2) 

p.RE
```

# Generate figure 7B
```{r}
res.metrics <- res %>% dplyr::select(Method, Dataset, ARI, NMI, `FM-Index`, FMeasure) %>% 
  as.data.frame() %>% 
  melt(id.var = c('Dataset','Method'))

colnames(res.metrics) <- c('Dataset', 'Method', 'Metric' ,'Score')

p.metrics <-ggpaired(res.metrics, x= 'Method', y = 'Score',
                 fill = 'Method', add = 'jitter', palette = 'jama',
                 point.size = 3, facet.by = 'Metric') + 
  labs(title="",
        x ="Cluster Estimation Method", y = "Score") + theme_bw() +
  theme(
      plot.title = element_text(color="Black", size=25, hjust = 0.5),
      plot.subtitle = element_text(color = "red", size = 25, face = "bold", hjust = 0.5),
      axis.title.x = element_text(color="Black", size=25),
      axis.title.y = element_text(color="Black", size=25),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 25),
      axis.text.y = element_text(size = 25),
      legend.text = element_text(size = 25),
      legend.title = element_text(size = 25),
    )

p.metrics

```

# Generate Figure 7
```{r, fig.height=8, fig.width=16}
cowplot::plot_grid(p.RE, p.metrics, labels = c('A','B'), rel_widths = c(2/5, 3/5))
```