---
title: "Distance Analysis"
output: html_document
date: '2022-08-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())
gc()
library(tidyverse)  # data manipulation
library(ggplot2)
library(data.table)
# library(rnndescent)
# library(imcdatasets)
library(flowAssist)
library(FlowSOM)
library(Rcpp)
library(parallel)
library(dplyr)
library(psych)
library(analogue)
library(ggstatsplot)
library(ggsci)
library(rcartocolor)
set.seed(1994)
library(miceadds)
source.all('~/Documents/PhD/Spatial_Clustering/clustering/scripts/')
```

# Maldegem2021
## read in the data and get markers of interest
```{r}
maldegem_dat <- as.data.frame(fread(
  '~/Documents/PhD/Spatial_Clustering/Maldegem2021_LungCancer_IMC/Maldegem2021_processed.csv'))

# remove the unclassified cell types
# maldegem_dat <- maldegem_dat[maldegem_dat$clustername3 != 'Unclassified', ]

# rename the cluster column
names(maldegem_dat)[names(maldegem_dat) == 'clustername3'] <- 'CellType'

maldegem_markers <- colnames(maldegem_dat)[1:17]

res.malde <- RunDistAnalysis(maldegem_dat, maldegem_markers)

rm(maldegem_dat)
gc()
```

# Rendeiro2020
## Read in the data and set markers
```{r}
rendeiro_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_intensities.csv',
             nThread = 32))

# get the metadata
rendeiro_meta <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_metadata.csv',
             nThread = 32))

# extract the markers of interest and add the cluster labels and image metatdata
rendeiro_markers <- colnames(rendeiro_dat)[1:38]
rendeiro_dat <- rendeiro_dat[, rendeiro_markers]

# combine the metadata with the intensities and umap coordinates
rendeiro_dat <- cbind(rendeiro_dat, cbind(rendeiro_meta))

names(rendeiro_dat)[names(rendeiro_dat) == 'metacluster_label'] <- 'CellType'
rendeiro_dat <- rendeiro_dat[rendeiro_dat$CellType != "nan", ]

res.ren <- RunDistAnalysis(rendeiro_dat, rendeiro_markers)

rm(rendeiro_dat)
gc()
```

# Risom2022
## Read in data and set markers
```{r}
risom_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Risom2022_BreastCancer_IMC/Single_Cell_Data.csv')
)

risom_markers <- c('CD45','SMA','CK7','CK5','VIM','CD31','PanKRT','ECAD',
                   'Tryptase','MPO','CD20','CD3','CD8','CD4','CD14','CD68','FAP',
                   'CD36','CD11c','HLADRDPDQ','P63','CD44')

names(risom_dat)[names(risom_dat) == 'manual_gating_phenotype'] <- 'CellType'

res.risom <- RunDistAnalysis(risom_dat, risom_markers)

rm(risom_dat)
gc()
```

# Damond2019 
# Read in data and set markers
```{r}
damond_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Damond2019_Diabetes_IMC/Damond2010_Diabetes_IMC_celldata.csv')
)


# get the markers of interest
damond_markers <- colnames(damond_dat)
damond_markers <- damond_markers[1:36]

res.damond <- RunDistAnalysis(damond_dat, damond_markers)

rm(damond_dat)
gc()
```

<!-- # Jackson2020 -->
<!-- ## Read in data and load markers -->
<!-- ```{r} -->
<!-- # get the normalized data -->
<!-- jackson_sce <- JacksonFischer2020Data(data_type = "sce") -->
<!-- jackson_dat <- as.data.frame(t(jackson_sce@assays@data$counts)) -->

<!-- # add celltypes to counts data -->
<!-- jackson_dat$CellType <- colData(jackson_sce)$metacluster -->

<!-- # exclude channels from publication -->
<!-- channel_exclude = c("ImageId" ,"CellId" ,"In115 115InIn115Di","Xe134 134XeXe134Di","Hg202 202HgHg202Di","Pb204 204PbPb204Di","Pb206 206PbPb206Di","ArAr80 80ArArArAr80Di","phospho Erk12", "10311239Ru96Di Rutheni","10311240Ru98Di Rutheni","10311241Ru99Di Rutheni", "10311242Ru100Di Rutheni","10311243Ru101Di Rutheni", "10311244Ru102Di Rutheni","10311245Ru104Di Rutheni","Xe126 126XeXe126Di","I127 127II127Di","Xe131 131XeXe131Di","Pb207 207PbPb207Di","Pb208 208PbPb208Di","EulerNumber","MajorAxisLength","MinorAxisLength", "Orientation","10331253Ir191Di Iridium","2971330Dy161Di EpCAM","Perimeter","1971527Ho165Di bCaten","483739Yb171Di Sox9","Solidity") -->

<!-- good_channels = unique(colnames(jackson_dat))[!unique(colnames(jackson_dat)) %in% channel_exclude] -->

<!-- jackson_dat <- jackson_dat[, good_channels] -->

<!-- # get the markers of interest -->
<!-- jackson_markers <- good_channels[1:34] -->

<!-- res.jackson <- RunDistAnalysis(jackson_dat, jackson_markers) -->

<!-- rm(jackson_dat) -->
<!-- gc() -->
<!-- ``` -->

# Moldoveanu2022
## read in data and set markers
```{r}
moldoveanu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/ICI_meanIntensity_processed.csv')
)

moldoveanu_dat_non_ICI <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/non-ICI_meanIntensity_processed.csv')
)


moldoveanu_metadata <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/ST4_cell_data.txt')
)

# combine both datasets
moldoveanu_dat <- rbind(moldoveanu_dat, moldoveanu_dat_non_ICI)

# order by object ID
moldoveanu_dat <- moldoveanu_dat[match(moldoveanu_metadata$obj.id, moldoveanu_dat$obj.id), ]

# get markers
moldoveanu_markers <- colnames(moldoveanu_dat)[1:12]

# add celltypes
moldoveanu_dat$CellType <- moldoveanu_metadata$Cluster

res.moldoveanu <- RunDistAnalysis(moldoveanu_dat, moldoveanu_markers)

rm(moldoveanu_dat)
gc()
```

# Hoch 2022
## read in data and set markers
```{r}
# read in the data
hoch_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Hoch2022_Melanoma_IMC/hoch2022_Melanoma_counts.csv', 
                  nThread = 7)
)
# get markers of interest
hoch_markers <- colnames(hoch_dat)[1:41]

res.hoch <- RunDistAnalysis(hoch_dat, hoch_markers)

rm(hoch_markers)
gc()
```

# Keren2018
## Read in data set markers
```{r}
keren_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Keren2018_BreastCancer_IMC/cellData.csv'))

# set markers
keren_markers <- c('CD45', 'FoxP3', 'CD4', 'CD8', 'CD3', 'CD20', 'CD16', 'CD68', 
                   'MPO', 'HLA-DR', 'Pan-Keratin', 'Keratin17', 'Keratin6', 'p53', 
                   'Beta catenin', 'EGFR')
# set the cell types
names(keren_dat)[names(keren_dat) == 'Group'] <- 'CellType'
```

## Run clustering
```{r}
res.keren <- RunDistAnalysis(keren_dat, keren_markers)

rm(keren_dat)
gc()
```

# McCaffery2022
# load data and get markers
```{r}
# read in the data
mccaferry_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/McCaffery2022_Tuberculosis_IMC/allTB-ME_annotated.csv')
)
# get the markers
mccaferry_markers <- colnames(mccaferry_dat)[6:42]

# set the cell types
names(mccaferry_dat)[names(mccaferry_dat) == 'cell_type'] <- 'CellType'

mccaferry_dat <- subset(mccaferry_dat,
                        select = c(mccaferry_markers, 'CellType'))

# remove rows with NAs
mccaferry_dat <- na.omit(mccaferry_dat)

res.mccaferry <- RunDistAnalysis(mccaferry_dat, mccaferry_markers)

rm(mccaferry_dat)
gc()
```

## Liu2022
# Get data and extract markers
```{r}
liu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Liu2022_Benchmarking_IMC/cell_table_size_normalized_clusters.csv')
)

liu_markers <- c('CD3', 'CD4', 'CD8', 'CD11c', 'CD20', 'CD31', 'CD45', 'CD56', 
                 'CD68', 'PANCK', 'PAX5','Vimentin')

# define cell types

names(liu_dat)[names(liu_dat) == 'phenotype'] <- 'CellType'

res.liu <- RunDistAnalysis(liu_dat, liu_markers)

rm(liu_dat)
gc()
```

## Schurch2020
# Get data and extract markers
```{r}
schurch_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/CODEX_Colon_Schurch2020/CRC_clusters_neighborhoods_markers.csv") %>%
    select(-1) %>% 
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(schurch_dat$cluster_name)
schurch_dat$CellType <- factor(schurch_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

# get the markers
schurch_markers <- colnames(schurch_dat)[12:60]

res.schurch <- RunDistAnalysis(schurch_dat, schurch_markers)

rm(schurch_dat)
gc()
```

## Phillips2021
# Get data and extract markers
```{r}
phillips_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/Phillips2021_Lymphoma_CODEX/Raw_df_CODEX.csv") %>%
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(phillips_dat$cluster_name)
phillips_dat$CellType <- factor(phillips_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

phillips_dat <- phillips_dat[ , -which(names(phillips_dat) %in% 
                                         c('cd11b','cd16','cd164','ccr4','ccr6',
                                      'egfr', 'p53'))]


# get the markers
phillips_markers <- colnames(phillips_dat)[10:61]

res.phillips <- RunDistAnalysis(phillips_dat, phillips_markers)

rm(phillips_dat)
gc()
```

## Lohoff2022
# Get data and extract markers
```{r}
lohoff_dat <- as.data.frame(as.matrix(t(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/exprs.Rds'))))
lohoff_metadata <- as.data.frame(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/metadata.Rds'))

# remove Xist from columns
lohoff_dat <- lohoff_dat[, -which(colnames(lohoff_dat) %in% 'Xist')]

# get the top 50 pcs
lohoff_pcs <- as.data.frame(gmodels::fast.prcomp(t(lohoff_dat), center = F)$rotation[, 1:50])

# add the celltype information
lohoff_pcs$CellType <- lohoff_metadata$celltype_mapped_refined

lohoff_markers <- colnames(lohoff_pcs)[1:50]

rm(lohoff_dat)
gc()
```

# Run Clustering
```{r}
res.lohoff <- RunDistAnalysis(lohoff_pcs, lohoff_markers)
```

# Bortolomeazzi2020
# Get data and extract markers
```{r}
bortolomeazzi_dat <- fread('~/Documents/PhD/Spatial_Clustering/IMC_ColorectalCancer_Bortolomeazzi2020/Single_Cell_Data.csv') %>% 
  as.data.frame()


# get markers, assume they use all for clustering
bortolomeazzi_markers <- colnames(bortolomeazzi_dat)[16:45]

# add the celltype information
names(bortolomeazzi_dat)[names(bortolomeazzi_dat) == 'Population'] <- 'CellType'


res.bortolomeazzi <- RunDistAnalysis(bortolomeazzi_dat, bortolomeazzi_markers)

rm(bortolomeazzi_dat)
gc()
```

# Brbic2022 - Intestine
# Get data and extract markers
```{r}
brbic_int_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/B004_training_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_int_markers <- colnames(brbic_int_dat)[2:49]

# add the celltype information
names(brbic_int_dat)[names(brbic_int_dat) == 'cell_type_A'] <- 'CellType'

res.brbic_int <- RunDistAnalysis(brbic_int_dat, brbic_int_markers)

rm(brbic_int_dat)
gc()
```

# Brbic2022 - Tonsil
# Get data and extract markers
```{r}
brbic_tons_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/BE_Tonsil_l3_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_tons_markers <- colnames(brbic_tons_dat)[2:45]

# add the celltype information
names(brbic_tons_dat)[names(brbic_tons_dat) == 'cell_type'] <- 'CellType'

res.brbic_tons <- RunDistAnalysis(brbic_tons_dat, brbic_tons_markers)

rm(brbic_tons_dat)
gc()
```

# compile the results
```{r}
res <- rbind(res.malde , res.ren, res.risom, res.damond, 
             res.keren,
             res.moldoveanu, res.liu, res.hoch, res.schurch, 
             res.phillips, res.mccaferry, res.lohoff, res.bortolomeazzi,
             res.brbic_int, res.brbic_tons) %>% as.data.frame()
res$ARI <- as.numeric(res$ARI)
res$NMI <- as.numeric(res$NMI)
res$`FM-Index` <- as.numeric(res$`FM-Index`)
res$FMeasure <- as.numeric(res$FMeasure)
res$Method <- as.character(res$Method)

write.csv(res, '~/Documents/PhD/Spatial_Clustering/clustering//Distance_Analysis_results.csv', row.names = F)
```

```{r}
res <- read.csv('~/Documents/PhD/Spatial_Clustering/clustering/Distance_Analysis_results.csv')

names(res)[names(res) == 'FM.Index'] <- 'FM-Index'
```


# plot results
```{r}
res.melted <- melt(res)
# All, P +S, C+S, P+C, P, C, S
res.melted$Method <- if_else(res.melted$Method == 'Cosine + Spearmman', 
                             'Cosine + Spearman', res.melted$Method)

res.melted$Method  <- factor(res.melted$Method , levels=c("Pearson", "Spearman", 
                                                          "Cosine","Euclidean",
                                                          "Pearson + Cosine",
                                                          "Pearson + Spearman", 
                                                          "Cosine + Spearman",
                                                          "Pearson + Euclidean",
                                                          "Spearson + Euclidean",
                                                          "Cosine + Euclidean",
                                                          "Pearson + Spearman + Cosine",
                                                          "Pearson + Spearman + Euclidean",
                                                          "Spearman + Cosine + Euclidean",
                                                          "Pearson + Cosine + Euclidean",
                                                          "All"
                                                          ))

t <- res.melted %>% group_by(Method, variable) %>% summarise(value = mean(value))

t %>%
  group_by(variable) %>%
  summarise(max = max(value)) -> t.2

left_join(t, t.2) %>%
  mutate(Method = Method, color = value == max) %>%
  filter(color == TRUE)-> t.3

p.fmeasure <- ggplot(t[t$variable %like% 'FMeasure', ], aes(x=Method, y=value)) + 
  geom_point(aes(size = 15)) + 
  # geom_point(data=t.3, aes(x = Method, y = value), 
  #                                        color = "red") +
  facet_wrap(~variable, scales='free_y') + 
  labs(title = "", x = "", y = "Score") + 
  ylim(0.50, 0.70) +
  theme_bw() + 
  theme(legend.position  = 'None',
      plot.title = element_text(color="Black", size=25, face="bold", hjust = 0.5),
      plot.subtitle = element_text(color = "red", size = 25, hjust = 0.5),
      axis.title.x = element_text(color="Black", size=25),
      axis.title.y = element_text(color="Black", size=25),
      axis.text.y = element_text(size = 25),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 25),
      strip.text.x = element_text(size = 25)
    )


```


# Generate significance plot
```{r}
res$Type <- if_else(res$Method == 'All', 'All', 'Other')

p.sig.ari <- ggbetweenstats(
  data  = res, 
  x     = Type,
  y     = ARI,
  xlab = "",
  title = "", 
  type = 'parametric', 
  palette          = "uniform_startrek",
  package          = "ggsci",
  pairwise.comparisons = T,
  results.subtitle = F,
  outlier.tagging = F,
  pairwise.display = 'all',
  centrality.plotting = F,
  plot.type = 'box'
  ) + 
  theme(axis.text.x = element_text(size = 11))

p.sig.nmi <- ggbetweenstats(
  data  = res,
  x     = Type,
  y     = NMI,
  xlab = "",
  title = "", 
  type = 'parametric', 
  palette          = "uniform_startrek",
  package          = "ggsci",
  pairwise.comparisons = T,
  results.subtitle = F,
  outlier.tagging = F,
  pairwise.display = 'all',
  centrality.plotting = F,
  plot.type = 'box') + 
  theme(axis.text.x = element_text(size = 11))

p.sig.fm <- ggbetweenstats(
  data  = res,
  x     = Type,
  y     = `FM-Index`,
  xlab = "",
  title = "", 
  type = 'parametric', 
  palette          = "uniform_startrek",
  package          = "ggsci",
  pairwise.comparisons = T,
  results.subtitle = F,
  outlier.tagging = F,
  pairwise.display = 'all',
  centrality.plotting = F,
  plot.type = 'box') + 
  theme(axis.text.x = element_text(size = 11))

p.sig.fmeasure <- ggbetweenstats(
  data  = res,
  x     = Type,
  y     = FMeasure,
  xlab = "",
  title = "", 
  type = 'parametric', 
  palette          = "uniform_startrek",
  package          = "ggsci",
  pairwise.comparisons = T,
  results.subtitle = F,
  outlier.tagging = F,
  pairwise.display = 'all',
  centrality.plotting = F,
  plot.type = 'box') + 
  theme(axis.text.x = element_text(size = 11))

p.sig.final <- cowplot::plot_grid(p.sig.ari, p.sig.nmi, p.sig.fm, p.sig.fmeasure)
```

