---
title: "Hierchichal Analysis"
output: html_document
date: '2022-06-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
library(tidyverse)
library(ggplot2)
library(psych)
library(analogue)
library(data.table)
library(DiffCorr)
library(coop)
library(miceadds)
library(trqwe)
library(ggstatsplot)
library(ggsci)
library(rcartocolor)
library(parallelDist)
library(parallel)
library(umap)
library(RColorBrewer)
library(amap)
library(SingleCellExperiment)
library(paletteer)
source.all('/home/ewillie/Documents/PhD/Papers/FuseSOM_Paper/helper_scripts')
```

Proximity Metrics: Euclidean, Manhattan, Maximum
Correlation Metrics: Pearson, Spearman, Cosine

```{r}
prox_metrics <- c('euclidean', 'manhattan', 'maximum')
corr_metrics <- c('pearson','spearman','cosine')
```

Dataset: Maldegem 
```{r}
maldegem_dat <- as.data.frame(fread(
  '~/Documents/PhD/Spatial_Clustering/Maldegem2021_LungCancer_IMC/Maldegem2021_processed.csv'))

# remove the unclassified cell types
# maldegem_dat <- maldegem_dat[maldegem_dat$clustername3 != 'Unclassified', ]

# rename the cluster column
names(maldegem_dat)[names(maldegem_dat) == 'clustername3'] <- 'CellType'

maldegem_markers <- colnames(maldegem_dat)[1:17]
```

```{r}
maldegem_subsets <- subsetMatrix(maldegem_dat, markers = maldegem_markers, rep = 5, nrows = 20000)
res.corr.maldegem <- lapply(maldegem_subsets, run_corr, markers = maldegem_markers, 
                              corr_metrics = corr_metrics) %>% bind_rows()
res.prox.maldegem <- lapply(maldegem_subsets, run_prox, markers = maldegem_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()
res.maldegem <- rbind(res.corr.maldegem, res.prox.maldegem)
res.maldegem$Dataset <- rep('Maldegem et al', nrow(res.maldegem))
```


# Rendeiro2020
## Read in the data and set markers
```{r}
rendeiro_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_intensities.csv',
             nThread = 32))

# get the metadata
rendeiro_meta <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_metadata.csv',
             nThread = 32))

# extract the markers of interest and add the cluster labels and image metatdata
rendeiro_markers <- colnames(rendeiro_dat)[1:38]
rendeiro_dat <- rendeiro_dat[, rendeiro_markers]

# combine the metadata with the intensities and umap coordinates
rendeiro_dat <- cbind(rendeiro_dat, cbind(rendeiro_meta))

names(rendeiro_dat)[names(rendeiro_dat) == 'metacluster_label'] <- 'CellType'
rendeiro_dat <- rendeiro_dat[rendeiro_dat$CellType != "nan", ]
```

```{r}
rendeiro_subsets <- subsetMatrix(rendeiro_dat, markers = rendeiro_markers, rep = 5, nrows = 20000)
res.corr.rendeiro <- lapply(rendeiro_subsets, run_corr, markers = rendeiro_markers, 
                              corr_metrics = corr_metrics) %>% bind_rows()
res.prox.rendeiro <- lapply(rendeiro_subsets, run_prox, markers = rendeiro_markers, 
                              prox_metrics = prox_metrics) %>% bind_rows()
res.rendeiro <- rbind(res.corr.rendeiro, res.prox.rendeiro)
res.rendeiro$Dataset <- rep('Rendeiro et al', nrow(res.rendeiro))
```


# Risom2022
## Read in data and set markers
```{r}
risom_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Risom2022_BreastCancer_IMC/Single_Cell_Data.csv')
)

risom_markers <- c('CD45','SMA','CK7','CK5','VIM','CD31','PanKRT','ECAD',
                   'Tryptase','MPO','CD20','CD3','CD8','CD4','CD14','CD68','FAP',
                   'CD36','CD11c','HLADRDPDQ','P63','CD44')

names(risom_dat)[names(risom_dat) == 'manual_gating_phenotype'] <- 'CellType'
```


```{r}

risom_subsets <- subsetMatrix(risom_dat, markers = risom_markers, rep = 5, nrows = 20000)
res.corr.risom <- lapply(risom_subsets, run_corr, markers = risom_markers, 
                              corr_metrics = corr_metrics) %>% bind_rows()
res.prox.risom <- lapply(risom_subsets, run_prox, markers = risom_markers, 
                              prox_metrics = prox_metrics) %>% bind_rows()
res.risom <- rbind(res.corr.risom, res.prox.risom)
res.risom$Dataset <- rep('Risom et al', nrow(res.risom))
```

# Damond2019 
# Read in data and set markers
```{r}
damond_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Damond2019_Diabetes_IMC/Damond2010_Diabetes_IMC_celldata.csv')
)


# get the markers of interest
damond_markers <- colnames(damond_dat)
damond_markers <- damond_markers[1:36]
```

```{r}
damond_subsets <- subsetMatrix(damond_dat, markers = damond_markers, rep = 5, nrows = 20000)
res.corr.damond <- lapply(damond_subsets, run_corr, markers = damond_markers, 
                              corr_metrics = corr_metrics) %>% bind_rows()
res.prox.damond <- lapply(damond_subsets, run_prox, markers = damond_markers, 
                              prox_metrics = prox_metrics) %>% bind_rows()
res.damond <- rbind(res.corr.damond, res.prox.damond)
res.damond$Dataset <- rep('Damond et al', nrow(res.damond))
```

# Moldoveanu2022
## read in data and set markers
```{r}
moldoveanu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/ICI_meanIntensity_processed.csv')
)

moldoveanu_dat_non_ICI <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/non-ICI_meanIntensity_processed.csv')
)


moldoveanu_metadata <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/ST4_cell_data.txt')
)

# combine both datasets
moldoveanu_dat <- rbind(moldoveanu_dat, moldoveanu_dat_non_ICI)

# order by object ID
moldoveanu_dat <- moldoveanu_dat[match(moldoveanu_metadata$obj.id, moldoveanu_dat$obj.id), ]

# get markers
moldoveanu_markers <- colnames(moldoveanu_dat)[1:12]

# add celltypes
moldoveanu_dat$CellType <- moldoveanu_metadata$Cluster
```

```{r}
moldoveanu_subsets <- subsetMatrix(moldoveanu_dat, markers = moldoveanu_markers, rep = 5, nrows = 20000)
res.corr.moldoveanu <- lapply(moldoveanu_subsets, run_corr, markers = moldoveanu_markers, 
                              corr_metrics = corr_metrics) %>% bind_rows()
res.prox.moldoveanu <- lapply(moldoveanu_subsets, run_prox, markers = moldoveanu_markers, 
                              prox_metrics = prox_metrics) %>% bind_rows()
res.moldoveanu <- rbind(res.corr.moldoveanu, res.prox.moldoveanu)
res.moldoveanu$Dataset <- rep('Moldoveanu et al', nrow(res.moldoveanu))
```

# Hoch 2022
## read in data and set markers
```{r}
# read in the data
hoch_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Hoch2022_Melanoma_IMC/hoch2022_Melanoma_counts.csv', 
                  nThread = 32)
)
# get markers of interest
hoch_markers <- colnames(hoch_dat)[1:41]
```

```{r}
hoch_subsets <- subsetMatrix(hoch_dat, markers = hoch_markers, rep = 5, nrows = 20000)
res.corr.hoch <- lapply(hoch_subsets, run_corr, markers = hoch_markers, 
                          corr_metrics = corr_metrics) %>% bind_rows()
res.prox.hoch <- lapply(hoch_subsets, run_prox, markers = hoch_markers, 
                          prox_metrics = prox_metrics) %>% bind_rows()

res.hoch <- rbind(res.corr.hoch, res.prox.hoch)
res.hoch$Dataset <- rep('Hoch et al', nrow(res.hoch))
```

# Keren2018
## Read in data set markers
```{r}
keren_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Keren2018_BreastCancer_IMC/cellData.csv'))

# set markers
keren_markers <- c('CD45', 'FoxP3', 'CD4', 'CD8', 'CD3', 'CD20', 'CD16', 'CD68', 
                   'MPO', 'HLA-DR', 'Pan-Keratin', 'Keratin17', 'Keratin6', 'p53', 
                   'Beta catenin', 'EGFR')
# set the cell types
names(keren_dat)[names(keren_dat) == 'Group'] <- 'CellType'
```

## Run clustering
```{r}
keren_subsets <- subsetMatrix(keren_dat, markers = keren_markers, rep = 5, nrows = 20000)
res.corr.keren <- lapply(keren_subsets, run_corr, markers = keren_markers, 
                          corr_metrics = corr_metrics) %>% bind_rows()
res.prox.keren <- lapply(keren_subsets, run_prox, markers = keren_markers, 
                          prox_metrics = prox_metrics) %>% bind_rows()

res.keren <- rbind(res.corr.keren, res.prox.keren)
res.keren$Dataset <- rep('Keren et al', nrow(res.keren))
```

# McCaffrey2022
# load data and get markers
```{r}
# read in the data
mccaferry_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/McCaffery2022_Tuberculosis_IMC/allTB-ME_annotated.csv')
)
# get the markers
mccaferry_markers <- colnames(mccaferry_dat)[6:42]

# set the cell types
names(mccaferry_dat)[names(mccaferry_dat) == 'cell_type'] <- 'CellType'

mccaferry_dat <- subset(mccaferry_dat,
                        select = c(mccaferry_markers, 'CellType'))

# remove rows with NAs
mccaferry_dat <- na.omit(mccaferry_dat)
```

```{r}
mccaferry_subsets <- subsetMatrix(mccaferry_dat, rep = 5, markers = mccaferry_markers, nrows = 20000)
res.corr.mccaferry <- lapply(mccaferry_subsets, run_corr, markers = mccaferry_markers, 
                               corr_metrics = corr_metrics) %>% bind_rows()
res.prox.mccaferry <- lapply(mccaferry_subsets, run_prox, markers = mccaferry_markers, 
                               prox_metrics = prox_metrics) %>% bind_rows()

res.mccaferry <- rbind(res.corr.mccaferry, res.prox.mccaferry)
res.mccaferry$Dataset <- rep('McCaffery et al', nrow(res.mccaferry))
```

## Liu2022
# Get data and extract markers
```{r}
liu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Liu2022_Benchmarking_IMC/cell_table_size_normalized_clusters.csv')
)

liu_markers <- c('CD3', 'CD4', 'CD8', 'CD11c', 'CD20', 'CD31', 'CD45', 'CD56', 
                 'CD68', 'PANCK', 'PAX5','Vimentin')

# define cell types

names(liu_dat)[names(liu_dat) == 'phenotype'] <- 'CellType'
```


```{r}
liu_subsets <- subsetMatrix(liu_dat, markers = liu_markers, rep = 5, nrows = 20000)
res.corr.liu <- lapply(liu_subsets, run_corr, markers = liu_markers, 
                         corr_metrics = corr_metrics) %>% bind_rows()
res.prox.liu <- lapply(liu_subsets, run_prox, markers = liu_markers, 
                         prox_metrics = prox_metrics) %>% bind_rows()

res.liu <- rbind(res.corr.liu, res.prox.liu)
res.liu$Dataset <- rep('Liu et al', nrow(res.liu))
```


## Schurch2020
# Get data and extract markers
```{r}
schurch_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/CODEX_Colon_Schurch2020/CRC_clusters_neighborhoods_markers.csv") %>%
    select(-1) %>% 
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(schurch_dat$cluster_name)
schurch_dat$CellType <- factor(schurch_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

# get the markers
schurch_markers <- colnames(schurch_dat)[12:60]
```

```{r}
schurch_subsets <- subsetMatrix(schurch_dat, markers = schurch_markers, rep = 5, nrows = 20000)
res.corr.schurch <- lapply(schurch_subsets, run_corr, markers = schurch_markers, 
                             corr_metrics = corr_metrics) %>% bind_rows()
res.prox.schurch <- lapply(schurch_subsets, run_prox, markers = schurch_markers, 
                             prox_metrics = prox_metrics) %>% bind_rows()
res.schurch <- rbind(res.corr.schurch, res.prox.schurch)
res.schurch$Dataset <- rep('Schurch et al', nrow(res.schurch))
```

## Phillips2021
# Get data and extract markers
```{r}
phillips_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/Phillips2021_Lymphoma_CODEX/Raw_df_CODEX.csv") %>%
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(phillips_dat$cluster_name)
phillips_dat$CellType <- factor(phillips_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

phillips_dat <- phillips_dat[ , -which(names(phillips_dat) %in% 
                                         c('cd11b','cd16','cd164','ccr4','ccr6',
                                      'egfr', 'p53'))]


# get the markers
phillips_markers <- colnames(phillips_dat)[10:61]
```

```{r}
phillips_subsets <- subsetMatrix(phillips_dat, markers = phillips_markers, rep = 5, nrows = 20000)
res.corr.phillips <- lapply(phillips_subsets, run_corr, markers = phillips_markers, 
                            corr_metrics = corr_metrics) %>% bind_rows()
res.prox.phillips <- lapply(phillips_subsets, run_prox, markers = phillips_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()

res.phillips <- rbind(res.corr.phillips, res.prox.phillips)
res.phillips$Dataset <- rep('Phillips et al', nrow(res.phillips))
```

## Lohoff2022
# Get data and extract markers
```{r}
lohoff_dat <- as.data.frame(as.matrix(t(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/exprs.Rds'))))
lohoff_metadata <- as.data.frame(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/metadata.Rds'))

# remove Xist from columns
lohoff_dat <- lohoff_dat[, -which(colnames(lohoff_dat) %in% 'Xist')]

# get the top 50 pcs
lohoff_pcs <- as.data.frame(gmodels::fast.prcomp(t(lohoff_dat), center = F)$rotation[, 1:50])

# add the celltype information
lohoff_pcs$CellType <- lohoff_metadata$celltype_mapped_refined

lohoff_markers <- colnames(lohoff_pcs)[1:50]

```

```{r}
lohoff_subsets <- subsetMatrix(lohoff_pcs, markers = lohoff_markers, rep = 5, nrows = 20000)
res.corr.lohoff <- lapply(lohoff_subsets, run_corr, markers = lohoff_markers, 
                            corr_metrics = corr_metrics) %>% bind_rows()
res.prox.lohoff <- lapply(lohoff_subsets, run_prox, markers = lohoff_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()

res.lohoff <- rbind(res.corr.lohoff, res.prox.lohoff)
res.lohoff$Dataset <- rep('Lohoff et al', nrow(res.lohoff))
```

# Bortolomeazzi2020
# Get data and extract markers
```{r}
bortolomeazzi_dat <- fread('~/Documents/PhD/Spatial_Clustering/IMC_ColorectalCancer_Bortolomeazzi2020/Single_Cell_Data.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
bortolomeazzi_markers <- colnames(bortolomeazzi_dat)[16:45]

# add the celltype information
names(bortolomeazzi_dat)[names(bortolomeazzi_dat) == 'Population'] <- 'CellType'
```

```{r}
bortolomeazzi_subsets <- subsetMatrix(bortolomeazzi_dat, markers = bortolomeazzi_markers,  rep = 5, nrows = 20000)
res.corr.bortolomeazzi <- lapply(bortolomeazzi_subsets, run_corr, markers = bortolomeazzi_markers, 
                            corr_metrics = corr_metrics) %>% bind_rows()
res.prox.bortolomeazzi <- lapply(bortolomeazzi_subsets, run_prox, markers = bortolomeazzi_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()

res.bortolomeazzi <- rbind(res.corr.bortolomeazzi, res.prox.bortolomeazzi)
res.bortolomeazzi$Dataset <- rep('Bortolomeazzi et al', nrow(res.bortolomeazzi))
```

# Brbic2022 - Intestine
# Get data and extract markers
```{r}
brbic_int_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/B004_training_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_int_markers <- colnames(brbic_int_dat)[2:49]

# add the celltype information
names(brbic_int_dat)[names(brbic_int_dat) == 'cell_type_A'] <- 'CellType'
```

```{r}
brbic_int_subsets <- subsetMatrix(brbic_int_dat, markers = brbic_int_markers, rep = 5, nrows = 20000)
res.corr.brbic_int <- lapply(brbic_int_subsets, run_corr, markers = brbic_int_markers, 
                            corr_metrics = corr_metrics) %>% bind_rows()
res.prox.brbic_int <- lapply(brbic_int_subsets, run_prox, markers = brbic_int_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()

res.brbic_int <- rbind(res.corr.brbic_int, res.prox.brbic_int)
res.brbic_int$Dataset <- rep('Brbic et al intestine', nrow(res.brbic_int))
```


# Brbic2022 - Tonsil
# Get data and extract markers
```{r}
brbic_tons_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/BE_Tonsil_l3_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_tons_markers <- colnames(brbic_tons_dat)[2:45]

# add the celltype information
names(brbic_tons_dat)[names(brbic_tons_dat) == 'cell_type'] <- 'CellType'
```

```{r}
brbic_tons_subsets <- subsetMatrix(brbic_tons_dat, markers = brbic_tons_markers, rep = 5, nrows = 20000)
res.corr.brbic_tons <- lapply(brbic_tons_subsets, run_corr, markers = brbic_tons_markers, 
                            corr_metrics = corr_metrics) %>% bind_rows()
res.prox.brbic_tons <- lapply(brbic_tons_subsets, run_prox, markers = brbic_tons_markers, 
                            prox_metrics = prox_metrics) %>% bind_rows()

res.brbic_tons <- rbind(res.corr.brbic_tons, res.prox.brbic_tons)
res.brbic_tons$Dataset <- rep('Brbic et al tonsil', nrow(res.brbic_tons))
```



# combine the results
```{r}
res.combined <- rbind(res.maldegem, res.rendeiro, res.risom, res.damond,
                      res.hoch, res.moldoveanu,res.keren, res.liu, res.mccaferry, 
                      res.schurch,res.phillips, res.lohoff, res.bortolomeazzi,
                      res.brbic_int, res.brbic_tons)

# write.csv(res.combined, '~/Documents/PhD/Spatial_Clustering/clustering/hclust_analysis_results_all_20k.csv',
#           row.names = F)

```


```{r}
res.combined<- read.csv('/home/ewillie/Documents/PhD/Papers/FuseSOM_Paper/result_data/hclust_analysis_results_all_20k.csv')
colnames(res.combined)[[3]] <- "FM-Index"
res.combined$Metric <- str_to_title(res.combined$Metric)

res.combined <- res.combined %>% group_by(Metric, Dataset) %>% summarise(ARI=mean(ARI), NMI=mean(NMI), 
                                                        `FM-Index`=mean(`FM-Index`),
                                                        FMeasure=mean(FMeasure))
res.combined$Dataset <- if_else(res.combined$Dataset == 'Mccaferry et al', 
                                'McCaffrey et al', res.combined$Dataset)

res.combined.melted <- melt(res.combined, id.vars = c('Metric', 'Dataset'))


res.combined.melted$Metric <- factor(res.combined.melted$Metric, levels = c("Pearson", "Spearman",
                                            "Cosine", "Euclidean",
                                            "Manhattan", "Maximum"))

```

```{r}
res.combined.melted.sub <- res.combined.melted[res.combined.melted$Metric %in% c('Pearson', 'Euclidean'), ]
res.combined.melted.sub$Metric <- if_else(res.combined.melted.sub$Metric %like% 'Pearson', 'Correlation', 'Euclidean')

p.bar.hclust <- ggplot(res.combined.melted.sub, aes(x = Dataset, y = value, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 1) +
  ylim(0, 1) + facet_wrap(~variable) + theme_bw() +
  ylab("Score") + theme(
      plot.title = element_text(color="Black", size=25, face="bold", hjust = 0.5),
      plot.subtitle = element_text(color = "red", size = 25, hjust = 0.5),
      axis.title.x = element_text(color="Black", size=25),
      axis.title.y = element_text(color="Black", size=25),
      axis.text.y = element_text(size = 25),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 25),
      strip.text.x = element_text(size = 25),
      legend.title = element_text(size=25), #change legend title font size
      legend.text = element_text(size=25) #change legend text font size)
    ) + scale_fill_jama() + ggtitle('Hierarchical Clustering')
```



# Generate Figure 2
```{r, fig.height=7, fig.width=9}
t <- res.combined %>% 
  group_by(Metric,Dataset) %>% 
  summarise(ARI=ave(ARI), NMI=ave(NMI), `FM-Index`=ave(`FM-Index`),
            FMeasure=ave(FMeasure)) %>% unique()
names(t)[names(t) == 'Metric'] <- 'Distance'

t$Distance <- factor(t$Distance, levels = c("Pearson", "Spearman",
                                            "Cosine", "Euclidean",
                                            "Manhattan", "Maximum"))

# get ranking of values
t.melted <- melt(t, id.vars = c('Distance', 'Dataset'))
ranked.data <- t.melted %>% group_by(variable, Dataset) %>% 
  summarise(Distance=Distance, Dataset=Dataset, Score=value, Rank = rank(value))

ranked.data$Rank <- as.factor(ranked.data$Rank)

ranked.data$Distance <- factor(ranked.data$Distance, levels = c("Pearson", "Spearman",
                                                                "Cosine", "Euclidean",
                                                                "Manhattan", "Maximum"))


ggplot(ranked.data, 
                   aes(x = Dataset, y = Distance)) +
  geom_point(aes(size=Rank, color=Rank), alpha=1) +
  facet_wrap(~variable) + 
  labs( x= "", y = "") +
  theme(
      axis.title.x = element_text(color="Black", size=18, face="bold"),
      axis.text.y = element_text(vjust = 0.5, hjust=1, 
                                 size = 18, color="Black"),
      panel.background = element_blank(),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 18),
      strip.text.x = element_text(size = 18),
      legend.title = element_text(size = 18),
      legend.text = element_text(size = 18),
      legend.key=element_blank()
    ) + scale_color_futurama()

```
# Generate Figure S1
```{r, fig.height=10, fig.width=12}
p.box.hclust <- ggplot(res.combined.melted, aes(x = Dataset, y = value, color = Metric)) +
  geom_boxplot(size = 0.8) + facet_wrap(~variable, scales = 'free') + 
  theme_bw() +
  ylab("Score") + theme(
      plot.title = element_text(color="Black", size=18, face="bold", hjust = 0.5),
      plot.subtitle = element_text(color = "red", size = 18, hjust = 0.5),
      axis.title.x = element_text(color="Black", size=18),
      axis.title.y = element_text(color="Black", size=18),
      axis.text.y = element_text(size = 18),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 18),
      strip.text.x = element_text(size = 18),
      legend.title = element_text(size=18), #change legend title font size
      legend.text = element_text(size=18) #change legend text font size)
    ) + scale_color_startrek() + ggtitle('')

p.box.hclust
```

# Generate figure 1
```{r, fig.height=6, fig.width=8}
sce <- SingleCellExperiment(assays = list(counts = t(mccaferry_dat[, mccaferry_markers])
),
colData = mccaferry_dat$CellType)

set.seed(1994)
nb.cols <- 16
mycolors <- colorRampPalette(brewer.pal(12, "Set3"))(nb.cols)
mycolors_new <- paletteer_d("ggsci::category20b_d3")

# generate UMAP
sce <- scater::runUMAP(sce,
                         subset_row = mccaferry_markers,
                         exprs_values = "counts", 
                         n_threads = 7,
                         metric = 'cosine',
                         verbose = T)

# umap_df <- res$layout %>%
#   as.data.frame()
# colnames(umap_df) <- c('UMAP 1', 'UMAP 2')

mccaferry_dat$CellType <- gsub("\\_"," ",mccaferry_dat$CellType)
# CellType <- gsub("\\_"," ",CellType)
mccaferry_dat$CellType <- str_to_title(mccaferry_dat$CellType)
# CellType <- str_to_title(CellType)

mccaferry_dat$CellType <- dplyr::recode(mccaferry_dat$CellType,
                                         'Endothelial' = 'Endothelial cells',
                                         'Cd16 Cd14 Mono' = 'Cd16/Cd14 Monocytes',
                                         'Cd8 T' = 'Cd8 T cells',
                                         'Cd4 T' = 'Cd4 T cells',
                                         'Cd14 Mono' = 'Cd14 Monocytes',
                                         'Neutrophil' = 'Neutrophils',
                                         'Fibroblast' = 'Fibroblasts',
                                         'B Cell' = 'B Cells',
                                         'Cd68 Mac' = 'Cd68 Macrophages',
                                         'Cd206 Mac' = 'Cd206 Macrophages',
                                         'Cd11c Dc/Mono' = 'Cd11c Dendritic cells/Monocytes',
                                         'Cd11b/C Cd206 Mac/Mono' = 'Cd11b/C Cd206 Macrophages/Monocytes',
                                         'Mast' = 'Mast cells',
                                         'Cd163 Mac' = 'Cd163 Macrophages',
                                         'Treg' = 'Regulatory T cells',
                                         'Epithelial' = 'Epithelial cells')

sce$`Cell Type` <- mccaferry_dat$CellType

gc()

umap_df <- reducedDim(sce, 'UMAP') %>%
    as.data.frame()
colnames(umap_df) <- c('UMAP 1', 'UMAP 2')

umap_df$`Cell Type` <- sce$`Cell Type`
```

```{r, fig.height=10, fig.width=16}
p.orig <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + 
  geom_point(aes(color=`Cell Type`), size = 3) + 
  theme_bw() + 
  ggtitle("Annotated cells (McCaffrey et al)") +
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 30, 
                                  hjust = 0.5, face = 'bold'),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25, face = 'bold')
        ) + 
  scale_color_manual(values = mycolors_new) + 
  guides(color=guide_legend(ncol = 1, 
                            override.aes = list(size=12)))

p.orig
```


```{r, fig.height=4, fig.width=6}
# cluster the dataset using the six metrics
dat <- mccaferry_dat[, mccaferry_markers]

k = length(unique(mccaferry_dat$CellType))

# pearson
d.pear <- cor2dist(cor(t(dat)))
gc()
pear.clusts <- cutree(fastcluster::hclust(as.dist(d.pear), method = 'complete'), k)
pear.clusts <- as.factor(pear.clusts)
pear.metrics <- ComputeMetrics(mccaferry_dat$CellType, pear.clusts)
p.pear <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + 
  geom_point(aes(color=pear.clusts), size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Pearson \nARI=0.22, NMI=0.28, FM-Index=0.33, FMeasure=0.46") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, 
                                  hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) + 
  scale_color_manual(values = mycolors)
rm(d.pear)
gc()
```

```{r, fig.height=8, fig.width=12}
p.pear
```

```{r}
# spearman
d.spear <- cor2dist(cor(t(dat), method = 'spearman'))
gc()
spear.clusts <- cutree(hclust(as.dist(d.spear), method = 'complete'), k)
spear.clusts <- as.factor(spear.clusts)
spear.metrics <- ComputeMetrics(mccaferry_dat$CellType, spear.clusts)
p.spear <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + geom_point(aes(color=spear.clusts),
                                                                     size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Spearman \nARI=0.15; NMI=0.23; FM-Index=0.29; FMeasure=0.41") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) +
  scale_color_manual(values = mycolors)
rm(d.spear)
gc()
```


```{r, fig.height=8, fig.width=12}
p.spear
```

```{r}
# cosine
d.cos <- cor2dist(coop::tcosine(dat))
gc()
cos.clusts <- cutree(hclust(as.dist(d.cos), method = 'complete'), k)
cos.clusts <- as.factor(cos.clusts)
cos.metrics <- ComputeMetrics(mccaferry_dat$CellType, cos.clusts)
p.cos <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + geom_point(aes(color=cos.clusts),
                                                                   size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Cosine \nARI=0.10; NMI=0.23; FM-Index=0.27; FMeasure=0.37") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) + 
  scale_color_manual(values = mycolors)
rm(d.cos)
gc()
```

```{r, fig.height=8, fig.width=12}
p.cos
```

```{r}
# euclidean
d.eucl <- parallelDist(as.matrix(dat))
eucl.clusts <- cutree(fastcluster::hclust(d.eucl, method = 'complete'), k)
eucl.clusts <- as.factor(eucl.clusts)
eucl.metrics <- ComputeMetrics(mccaferry_dat$CellType, eucl.clusts)
p.eucl <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + 
  geom_point(aes(color=eucl.clusts), size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Euclidean \nARI=0.11, NMI=0.18, FM-Index=0.29, FMeasure=0.36") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) + 
  scale_color_manual(values = mycolors)
```

```{r, fig.height=8, fig.width=12}
p.eucl
```

```{r}
# manhattan
d.man <- parallelDist(as.matrix(dat), method = 'manhattan')
man.clusts <- cutree(fastcluster::hclust(d.man, method = 'complete'), k)
man.clusts <- as.factor(man.clusts)
man.metrics <- ComputeMetrics(mccaferry_dat$CellType, man.clusts)
p.man <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + geom_point(aes(color=man.clusts),
                                                                   size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Manhattan \nARI=0.10; NMI=0.21; FM-Index=0.22; FMeasure=0.33") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) + 
  scale_color_manual(values = mycolors)
```

```{r, fig.height=8, fig.width=12}
p.man
```

```{r}
# maximum
d.max <- parallelDist(as.matrix(dat), method = 'maximum')
max.clusts <- cutree(fastcluster::hclust(d.max, method = 'complete'), k)
max.clusts <- as.factor(max.clusts)
max.metrics <- ComputeMetrics(mccaferry_dat$CellType, max.clusts)
p.max <- ggplot(umap_df, aes(x=`UMAP 1`, y=`UMAP 2`)) + 
  geom_point(aes(color=max.clusts), size = 3) + 
  theme_bw() + theme(legend.position="none") + 
  ggtitle("Hierarchical clustering using Maximum \nARI=0.01; NMI=0.03; FM-Index=0.33; FMeasure=0.31") + 
  theme(axis.text.x = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        legend.text = element_text(size = 25),
        legend.title = element_text(size = 25)) + 
  scale_color_manual(values = mycolors)
```

```{r, fig.height=8, fig.width=12}
p.max
```

```{r, fig.height=18, fig.width=28}
cowplot::plot_grid(p.pear,p.cos, p.spear, p.eucl, p.man, p.max)
```

```{r}
gridExtra::grid.arrange(grid::nullGrob(), p.orig, grid::nullGrob(), p.pear, p.cos,
                        p.spear, byrow = TRUE, ncol = 3)
```




