---
title: "Pipeline Analysis Full"
output: html_document
date: '2022-08-17'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())
gc()
library(tidyverse)  # data manipulation
library(ggplot2)
library(data.table)
library(flowAssist)
library(FlowSOM)
library(Rcpp)
library(parallel)
library(dplyr)
library(psych)
library(analogue)
library(ggstatsplot)
library(ggsci)
library(rcartocolor)
library(FuseSOM)
set.seed(1994)
library(miceadds)
library(cowplot)
library(scDHA)
source.all('/home/ewillie/Documents/PhD/Papers/FuseSOM_Paper/helper_scripts')
```

# Maldegem2021
## read in the data and get markers of interest
```{r}
maldegem_dat <- as.data.frame(fread(
  '~/Documents/PhD/Spatial_Clustering/Maldegem2021_LungCancer_IMC/Maldegem2021_processed.csv'))

# remove the unclassified cell types
# maldegem_dat <- maldegem_dat[maldegem_dat$clustername3 != 'Unclassified', ]

# rename the cluster column
names(maldegem_dat)[names(maldegem_dat) == 'clustername3'] <- 'CellType'

maldegem_markers <- colnames(maldegem_dat)[1:17]

res.malde.full <- RunPipelineFull(maldegem_dat, maldegem_markers)
# get the mean values
res.malde.full <- setDT(res.malde.full)[,list(ARI=mean(ARI), NMI=mean(NMI),
                                              `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.malde.full$Dataset <- rep('Maldegem et al', 2)

rm(maldegem_dat)
gc()
```


```{r}
#Generate clustering result, the input matrix has rows as samples and columns as genes
result <- scDHA(as.matrix(maldegem_dat[, maldegem_markers]), 
                seed = 1994, ncores = 20)

#The clustering result can be found here 
cluster <- result$cluster
```



# Rendeiro2020
## Read in the data and set markers
```{r}
rendeiro_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_intensities.csv',
             nThread = 32))

# get the metadata
rendeiro_meta <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Rendeiro2020_Covid_IMC/Rendeiro2020_Covid_IMC_metadata.csv',
             nThread = 32))

# extract the markers of interest and add the cluster labels and image metatdata
rendeiro_markers <- colnames(rendeiro_dat)[1:38]
rendeiro_dat <- rendeiro_dat[, rendeiro_markers]

# combine the metadata with the intensities and umap coordinates
rendeiro_dat <- cbind(rendeiro_dat, cbind(rendeiro_meta))

names(rendeiro_dat)[names(rendeiro_dat) == 'metacluster_label'] <- 'CellType'
rendeiro_dat <- rendeiro_dat[rendeiro_dat$CellType != "nan", ]

res.ren.full <- RunPipelineFull(rendeiro_dat, rendeiro_markers)
# get the mean values
res.ren.full <- setDT(res.ren.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.ren.full$Dataset <- rep('Rendeiro et al', 2)

rm(rendeiro_dat)
gc()
```

# Risom2022
## Read in data and set markers
```{r}
risom_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Risom2022_BreastCancer_IMC/Single_Cell_Data.csv')
)

risom_markers <- c('CD45','SMA','CK7','CK5','VIM','CD31','PanKRT','ECAD',
                   'Tryptase','MPO','CD20','CD3','CD8','CD4','CD14','CD68','FAP',
                   'CD36','CD11c','HLADRDPDQ','P63','CD44')

names(risom_dat)[names(risom_dat) == 'manual_gating_phenotype'] <- 'CellType'

res.risom.full <- RunPipelineFull(risom_dat, risom_markers)
# get the mean values
res.risom.full <- setDT(res.risom.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.risom.full$Dataset <- rep('Risom et al', 2)

rm(risom_dat)
gc()
```

# Damond2019 
# Read in data and set markers
```{r}
damond_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Damond2019_Diabetes_IMC/Damond2010_Diabetes_IMC_celldata.csv')
)


# get the markers of interest
damond_markers <- colnames(damond_dat)
damond_markers <- damond_markers[1:36]

res.damond.full <- RunPipelineFull(damond_dat, damond_markers)
# get the mean values
res.damond.full <- setDT(res.damond.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.damond.full$Dataset <- rep('Damond et al', 2)

rm(damond_dat)
gc()
```


# Moldoveanu2022
## read in data and set markers
```{r}
moldoveanu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/data/ICI_meanIntensity_processed.csv')
)

moldoveanu_dat_non_ICI <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/non-ICI_meanIntensity_processed.csv')
)


moldoveanu_metadata <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Moldoveanu20222_Melanoma_IMC/ST4_cell_data.txt')
)

# combine both datasets
moldoveanu_dat <- rbind(moldoveanu_dat, moldoveanu_dat_non_ICI)

# order by object ID
moldoveanu_dat <- moldoveanu_dat[match(moldoveanu_metadata$obj.id, moldoveanu_dat$obj.id), ]

# get markers
moldoveanu_markers <- colnames(moldoveanu_dat)[1:12]

# add celltypes
moldoveanu_dat$CellType <- moldoveanu_metadata$Cluster

res.moldoveanu.full <- RunPipelineFull(moldoveanu_dat, moldoveanu_markers)
# get the mean values
res.moldoveanu.full <- setDT(res.moldoveanu.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.moldoveanu.full$Dataset <- rep('Moldoveanu et al', 2)

rm(moldoveanu_dat)
gc()
```

# Hoch 2022
## read in data and set markers
```{r}
# read in the data
hoch_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Hoch2022_Melanoma_IMC/hoch2022_Melanoma_counts.csv', 
                  nThread = 7)
)
# get markers of interest
hoch_markers <- colnames(hoch_dat)[1:41]

res.hoch.full <- RunPipelineFull(hoch_dat, hoch_markers)
# get the mean values
res.hoch.full <- setDT(res.hoch.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.hoch.full$Dataset <- rep('Hoch et al', 2)

rm(hoch_dat)
gc()
```

# Keren2018
## Read in data set markers
```{r}
keren_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Keren2018_BreastCancer_IMC/cellData.csv'))

# set markers
keren_markers <- c('CD45', 'FoxP3', 'CD4', 'CD8', 'CD3', 'CD20', 'CD16', 'CD68', 
                   'MPO', 'HLA-DR', 'Pan-Keratin', 'Keratin17', 'Keratin6', 'p53', 
                   'Beta catenin', 'EGFR')
# set the cell types
names(keren_dat)[names(keren_dat) == 'Group'] <- 'CellType'

res.keren.full <- RunPipelineFull(keren_dat, keren_markers)
# get the mean values
res.keren.full <- setDT(res.keren.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.keren.full$Dataset <- rep('Keren et al', 2)

rm(keren_dat)
gc()
```


# McCaffery2022
# load data and get markers
```{r}
# read in the data
mccaferry_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/McCaffery2022_Tuberculosis_IMC/allTB-ME_annotated.csv')
)
# get the markers
mccaferry_markers <- colnames(mccaferry_dat)[6:42]

# set the cell types
names(mccaferry_dat)[names(mccaferry_dat) == 'cell_type'] <- 'CellType'

mccaferry_dat <- subset(mccaferry_dat,
                        select = c(mccaferry_markers, 'CellType'))

# remove rows with NAs
mccaferry_dat <- na.omit(mccaferry_dat)

res.mccaferry.full <- RunPipelineFull(mccaferry_dat, mccaferry_markers)
# get the mean values
res.mccaferry.full <- setDT(res.mccaferry.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]

res.mccaferry.full$Dataset <- rep('Mccaferry et al', 2)

rm(mccaferry_dat)
gc()
```

## Liu2022
# Get data and extract markers
```{r}
liu_dat <- as.data.frame(
  fread('~/Documents/PhD/Spatial_Clustering/Liu2022_Benchmarking_IMC/cell_table_size_normalized_clusters.csv')
)

liu_markers <- c('CD3', 'CD4', 'CD8', 'CD11c', 'CD20', 'CD31', 'CD45', 'CD56', 
                 'CD68', 'PANCK', 'PAX5','Vimentin')

# define cell types

names(liu_dat)[names(liu_dat) == 'phenotype'] <- 'CellType'

res.liu.full <- RunPipelineFull(liu_dat, liu_markers)
# get the mean values
res.liu.full <- setDT(res.liu.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]

res.liu.full$Dataset <- rep('Liu et al', 2)

rm(liu_dat)
gc()
```

## Schurch2020
# Get data and extract markers
```{r}
schurch_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/CODEX_Colon_Schurch2020/CRC_clusters_neighborhoods_markers.csv") %>%
    select(-1) %>% 
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(schurch_dat$cluster_name)
schurch_dat$CellType <- factor(schurch_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

# get the markers
schurch_markers <- colnames(schurch_dat)[12:60]

res.schurch.full <- RunPipelineFull(schurch_dat, schurch_markers)
# get the mean values
res.schurch.full <- setDT(res.schurch.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]

res.schurch.full$Dataset <- rep('Schurch et al', 2)

rm(schurch_dat)
gc()
```

## Phillips2021
# Get data and extract markers
```{r}
phillips_dat <- read_csv("~/Documents/PhD/Spatial_Clustering/Phillips2021_Lymphoma_CODEX/Raw_df_CODEX.csv") %>%
    janitor::clean_names() %>% 
  as.data.frame() 

lev <- unique(phillips_dat$cluster_name)
phillips_dat$CellType <- factor(phillips_dat$cluster_name, levels = lev, labels = janitor::make_clean_names(lev))

phillips_dat <- phillips_dat[ , -which(names(phillips_dat) %in% 
                                         c('cd11b','cd16','cd164','ccr4','ccr6',
                                      'egfr', 'p53'))]


# get the markers
phillips_markers <- colnames(phillips_dat)[10:61]

res.phillips.full <- RunPipelineFull(phillips_dat, phillips_markers)
# get the mean values
res.phillips.full <- setDT(res.phillips.full)[,list(ARI=mean(ARI), NMI=mean(NMI), `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.phillips.full$Dataset <- rep('Phillips et al', 2)

rm(phillips_dat)
gc()
```

## Lohoff2022
# Get data and extract markers
```{r}
lohoff_dat <- as.data.frame(as.matrix(t(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/exprs.Rds'))))
lohoff_metadata <- as.data.frame(readRDS('~/Documents/PhD/Spatial_Clustering/Lohoff2022_Mouse_seqFISH/metadata.Rds'))

# remove Xist from columns
lohoff_dat <- lohoff_dat[, -which(colnames(lohoff_dat) %in% 'Xist')]

# get the top 50 pcs
lohoff_pcs <- as.data.frame(gmodels::fast.prcomp(t(lohoff_dat), center = F)$rotation[, 1:50])

# add the celltype information
lohoff_pcs$CellType <- lohoff_metadata$celltype_mapped_refined

lohoff_markers <- colnames(lohoff_pcs)[1:50]

res.lohoff.full <- RunPipelineFull(lohoff_pcs, lohoff_markers)
# get the mean values
res.lohoff.full <- setDT(res.lohoff.full)[,list(ARI=mean(ARI), NMI=mean(NMI),
                                                    `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.lohoff.full$Dataset <- rep('Lohoff et al', 2)

rm(lohoff_dat)
gc()
```

# Bortolomeazzi2020
# Get data and extract markers
```{r}
bortolomeazzi_dat <- fread('~/Documents/PhD/Spatial_Clustering/IMC_ColorectalCancer_Bortolomeazzi2020/Single_Cell_Data.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
bortolomeazzi_markers <- colnames(bortolomeazzi_dat)[16:45]

# add the celltype information
names(bortolomeazzi_dat)[names(bortolomeazzi_dat) == 'Population'] <- 'CellType'


res.bortolomeazzi.full <- RunPipelineFull(bortolomeazzi_dat, bortolomeazzi_markers)
# get the mean values
res.bortolomeazzi.full <- setDT(res.bortolomeazzi.full)[,list(ARI=mean(ARI), NMI=mean(NMI),
                                                    `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.bortolomeazzi.full$Dataset <- rep('Bortolomeazzi et al', 2)

rm(bortolomeazzi_dat)
gc()
```

# Brbic2022 - Intestine
# Get data and extract markers
```{r}
brbic_int_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/B004_training_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_int_markers <- colnames(brbic_int_dat)[2:49]

# add the celltype information
names(brbic_int_dat)[names(brbic_int_dat) == 'cell_type_A'] <- 'CellType'

res.brbic_int.full <- RunPipelineFull(brbic_int_dat, brbic_int_markers)
# get the mean values
res.brbic_int.full <- setDT(res.brbic_int.full)[,list(ARI=mean(ARI), NMI=mean(NMI),
                                                    `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),  
                                                by=c('Method')]
res.brbic_int.full$Dataset <- rep('Brbic et al intestine', 2)

rm(brbic_int_dat)
gc()
```

# Brbic2022 - Tonsil
# Get data and extract markers
```{r}
brbic_tons_dat <- fread('~/Documents/PhD/Spatial_Clustering/CODEX_Intestine_Tonsil_Brbic2022/BE_Tonsil_l3_dryad.csv') %>% 
  as.data.frame()

# get markers, assume they use all for clustering
brbic_tons_markers <- colnames(brbic_tons_dat)[2:45]

# add the celltype information
names(brbic_tons_dat)[names(brbic_tons_dat) == 'cell_type'] <- 'CellType'

res.brbic_tons.full <- RunPipelineFull(brbic_tons_dat, brbic_tons_markers)
# get the mean values
res.brbic_tons.full <- setDT(res.brbic_tons.full)[,list(ARI=mean(ARI), NMI=mean(NMI),
                                                    `FM-Index`=mean(`FM-Index`), FMeasure = mean(FMeasure)),
                              by=c('Method')]
res.brbic_tons.full$Dataset <- rep('Brbic et al tonsil', 2)

rm(brbic_tons_dat)
gc()
```

```{r}
res <- rbind(res.malde.full , res.ren.full, res.risom.full, res.damond.full, 
             res.keren.full,
             res.moldoveanu.full, res.liu.full, res.hoch.full, res.schurch.full, 
             res.phillips.full, res.mccaferry.full, res.lohoff.full, res.bortolomeazzi.full,
             res.brbic_int.full, res.brbic_tons.full) %>% as.data.frame()

# write.csv(res, '~/Documents/PhD/Spatial_Clustering/clustering/FuseSOM_vs_FlowSOM_correlation.csv',
#           row.names = F)
```

```{r}
res <- read.csv('~/Documents/PhD/Spatial_Clustering/clustering/FuseSOM_vs_FlowSOM_Vanilla_updated.csv',
                check.names = FALSE)

res.melted <- melt(res)
res.melted$Method <- factor(res.melted$Method, 
                           levels = c('FuseSOM', 'FlowSOM'))
```

# Generate Figure 6B
```{r}
p.box.all <- ggpaired(res.melted, x = "Method", y = "value",
          color = "Method", palette = "startrek", 
          add = "jitter", line.color = "gray", 
          line.size = 1, point.size = 3, facet.by = 'variable') +
  labs(title = "", x = "", y = "ARI") + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20),
        strip.text.x = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        plot.subtitle = element_text(size = 20)) + 
  stat_compare_means(paired = TRUE, size = 6,
                     label.x = 1.5, label = "p.format") +
  geom_line(aes(group = Dataset), alpha = 0.1)

```


# Generate Figure 6 A
```{r}
res$Method <- factor(res$Method, levels = c('FuseSOM', 'FlowSOM'))
p.bar.ari <- ggplot(res, aes(x=Dataset, y=ARI, fill=Method)) + 
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 1) + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 20),
        axis.text.y = element_text(size = 20),
        plot.title = element_text(color="Black", size=25,hjust = 0.5),
        legend.position = 'None') + 
  labs(title = "ARI", x = "", y = "") + scale_fill_startrek()

p.bar.nmi <- ggplot(res, aes(x=Dataset, y=NMI, fill=Method)) + 
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 1) + 
  theme_bw() + theme(legend.position="none") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 20),
        plot.title = element_text(color="Black", size=20,hjust = 0.5),
        axis.text.y = element_text(size = 20)) + 
  labs(title = "NMI", x = "", y = "") + 
  coord_flip() + scale_fill_startrek()

p.bar.fm <- ggplot(res, aes(x=Dataset, y=`FM-Index`, fill=Method)) + 
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 1) + 
  theme_bw() + theme(legend.position="none") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 20),
        plot.title = element_text(color="Black", size=20,hjust = 0.5),
        axis.text.y = element_text(size = 20)) + 
  labs(title = "FM-Index", x = "", y = "") + 
  coord_flip() + scale_fill_startrek()

p.bar.fmeasure <- ggplot(res, aes(x=Dataset, y=FMeasure, fill=Method)) + 
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 1) + 
  theme_bw() + theme(legend.position="none") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 20),
        plot.title = element_text(color="Black", size=20,hjust = 0.5),
        axis.text.y = element_text(size = 20)) + 
  labs(title = "Fmeasure", x = "", y = "") + 
  coord_flip() + scale_fill_startrek()
```

```{r}
p.bar.final <- plot_grid(p.bar.ari, p.bar.nmi, p.bar.fm, p.bar.fmeasure)
```

# Generate Figure 6
```{r, fig.height=10, fig.width=18}
p.final <- plot_grid(p.bar.final, p.box.all, labels = c('A', 'B'), rel_widths = c(5/10, 5/10))
p.final
```
